---
title: "Skills Quiz: Residuals, Sums of Squares, and R-squared"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
    toc: true
    toc_float: true
---


## Instructions

Use this file to keep a record of your work as you complete the "Skills Quiz: Residuals, Sums of Squares, and R-squared" assignment in Canvas.


----

<!-- Note: The {} after each Problem and Part header allows you to keep track of what work you have completed. Write something like {Done} once you complete each problem and your html file will then show you a nice summary of what you have "done" already. -->

## Problem 1 {done}

Open the `Orange` dataset in R. As stated in the help file for this data set, "The Orange data... records of the growth of orange trees." 

Perform a simple linear regression of the circumference of the tree based on its age in days.

```{r}
lm.orange <- lm(circumference ~ age, data = Orange)
summary(lm.orange)
```

### Part (a) {done}

Type out the mathematical equation for this regression model and label both $Y$ and $X$ in the equation.

<div class="YourAnswer">

$$
  \underbrace{Y_i}_\text{circumference} = \beta_0 + \beta_1 \underbrace{X_i}_\text{age} + \epsilon_i \quad \text{where} \ \epsilon_i \sim N(0,\sigma^2) 
$$

</div>


### Part (b) {done}
 
Plot a scatterplot of the data with your regression line overlaid.

<div class="YourAnswer">

```{r}
Orange %>% 
  ggplot(aes(age, circumference)) +
  geom_point(aes(), color = "darkorange2") +
  geom_smooth(method = "lm", se = F, color = "firebrick") +
  labs(title = "Growth of Orange Trees",
       subtitle = "Orange data set",
       y = "Circumference of Tree in mm (circumference)",
       x = "Age of Tree in Days (age)") +
  theme(panel.background = element_blank(),
        panel.grid = element_blank())

```

</div>


### Part (c) {done}

State and interpret the slope and y-intercept of this regression. Are they meaningful for this data?

<div class="YourAnswer">

**Slope** <br>
The average `circumefence` increases in $0.11$ (mm) for each day increasing in `age` <br>

<br>
**Y-intercept** <br>
The average `circumference` is $17.4$ when `age` is $0$, in other words, when the tree is planted.

</div> 
 
 
### Part (d) {done}

What are the values of SSE, SSR, SSTO, $R^2$, and the correlation $r$ for this regression?
 
<div class="YourAnswer">

```{r}
sse <- sum((Orange$circumference - lm.orange$fit) ^ 2)
sse

ssr <- sum((lm.orange$fit - mean(Orange$circumference)) ^ 2)
ssr

ssto <- sum((Orange$circumference - mean(Orange$circumference)) ^ 2)
ssto

r2 <- ssr/ssto
r2

r <- sqrt(r2)
r
```

</div>


### Part (e) {done}

How do the values of SSR and SSTO relate to the correlation?

<div class="YourAnswer">

The `SSR` divided by `SSTO` gives the proportion of variation in `Y` explained by the regression, which is R-squared. 

</div>


### Part (f) {}

What circumference would we expect orange trees to have on average after 3 years, based on this regression? 

<div class="YourAnswer">

```{r}
predict(lm.orange, data.frame(age = 365), interval = "prediction")[1]
```

</div>



----

## Problem 2 {done}

Open the **mtcars** dataset in R. Fit three different regression models to the data that can each be used to explain average gas mileage of a vehicle (`mpg`). 

* The first regression should use the weight (`wt`) of the vehicle as the explanatory variable.
* The second should use the number of cylinders (`cyl`) of the engine of the vehicle as the explanatory variable.
* The third should use the gross horsepower of the vehicle (`hp`) as the explanatory variable.

### Part (a) {done}

Plot all three regressions in three separate plots.

<div class="YourAnswer">

```{r}
plot.f <- function(x, y, title) {
  p <- mtcars %>% 
    ggplot(aes(x, y)) +
    geom_point(aes(), color = "lightsteelblue") +
    geom_smooth(method = "lm", se = F, color = "steelblue") +
    labs(title = title) +
    theme(panel.background = element_blank(),
        panel.grid = element_blank())
p
}

plot.f(mtcars$wt, mtcars$mpg, "Weight")
plot.f(mtcars$cyl, mtcars$mpg, "Cylinders")
plot.f(mtcars$hp, mtcars$mpg, "Horse Power")


```

</div>


### Part (b) {done}

State the values of SSE, SSR, SSTO, and $R^2$ for each regression. 

<div class="YourAnswer">

##### `mpg ~ wt`

```{r}
lm.wt <- lm(mpg ~ wt, data = mtcars)
```

**SSE** <br>
```{r}
sse <- round(sum((mtcars$mpg - lm.wt$fit) ^2), 2)
sse
```

**SSR** <br>
```{r}
ssr <- round(sum((lm.wt$fit - mean(mtcars$mpg)) ^2), 2)
ssr
```

**SSTO** <br>
```{r}
ssto <- round(sum((mtcars$mpg - mean(mtcars$mpg)) ^2), 2)
ssto
```

**R^2** <br>
```{r}
r2 <- round(ssr / ssto, 2)
r2
```


##### `mpg ~ cyl`

```{r}
lm.cyl <- lm(mpg ~ cyl, data = mtcars)
```

**SSE** <br>
```{r}
sse <- round(sum((mtcars$mpg - lm.cyl$fit) ^2), 2)
sse
```

**SSR** <br>
```{r}
ssr <- round(sum((lm.cyl$fit - mean(mtcars$mpg)) ^2), 2)
ssr
```

**SSTO** <br>
```{r}
ssto <- round(sum((mtcars$mpg - mean(mtcars$mpg)) ^2), 2)
ssto
```

**R^2** <br>
```{r}
r2 <- round(ssr / ssto, 2)
r2
```




##### `mpg ~ hp`

```{r}
lm.hp <- lm(mpg ~ hp, data = mtcars)
```

**SSE** <br>
```{r}
sse <- round(sum((mtcars$mpg - lm.hp$fit) ^2), 2)
sse
```

**SSR** <br>
```{r}
ssr <- round(sum((lm.hp$fit - mean(mtcars$mpg)) ^2), 2)
ssr
```

**SSTO** <br>
```{r}
ssto <- round(sum((mtcars$mpg - mean(mtcars$mpg)) ^2), 2)
ssto
```

**R^2** <br>
```{r}
r2 <- round(ssr / ssto, 2)
r2
```






</div>


### Part (c) {done}

Compare the values from **Part (b)** across each regression. Consider these questions as you compare these numbers.

* What insight do these numbers give about the regression?
* Which numbers change, which stay the same, in these regressions? Why?
* Which regression is best at explaining average `mpg` according to these numbers?

<div class="YourAnswer">

`wt`, because has the lowest SSE and highest correlation.

</div> 
 
 



----

Before we can really trust the $R^2$ value from a regression model, there are important diagnostic checks to perform on the regression. 

You now have a better understanding of what a **residual** and a **fitted-value** are in regression. With that improved knowledge, work through the following problem.



----

## Problem 3 {}

Check the following technical details for each of the three regressions of **Problem 2**.

### Part (a) {}

Create a (1) residuals vs. fitted-values plot, (2) Q-Q Plot of the residuals, and (3) residuals vs. order plot for each regression of **Problem 2**.

<div class="YourAnswer">

```{r}
# Type your code here
```

</div>


### Part (b) {}

Explain, as best you understand currently, what each of these three plots show for these regressions.

<div class="YourAnswer">

Type your answer here...

</div>


### Part (c) {}

Comment on whether or not we should trust the $R^2$ value from each of your three regressions based on your plots in Part (a). 

<div class="YourAnswer">

Type your answer here...

</div>



----











<style>

.YourAnswer {
  color: #317eac;
  padding: 10px;
  border-style: solid;
  border-width: 2px;
  border-color: skyblue4;
  border-radius: 5px;
}

</style>

 
 